---
title: "Interactive Crime Map - Writeup"
author: "Scott Robertson"
date: "18/02/2019"
output: html_document
---

```{r setup}
# Set working directory
setwd("~/R Programming/Coursera/Data Products/Shiny/LondonCrimeMap")

# Load required packages
if("dplyr" %in% rownames(installed.packages()) == FALSE){
      install.packages("dplyr")
}
if("tidyr" %in% rownames(installed.packages()) == FALSE){
      install.packages("tidyr")
}
```

## Synopsis

The purpose of this paper is to document the development of an interactive crime map for London using open data from [data.police.uk](https://data.police.uk).

Inspiration and initial code were provided by two articles:

- https://towardsdatascience.com/plotting-a-map-of-london-crime-data-using-r-8dcefef1c397
- https://thisisnicblog.wordpress.com/

The final product will be presented a shiny dashboard.


## Data

Street level crime data can be accessed from [data.police.uk](https://data.police.uk) which holds individual crime data for all UK police forces going back to 2013. An API interface is also avaliable, however as I am not particularly good with APIs I have not used it for this project.

To build the dashboard I have downloaded csv files for the period January 2016 - Decemeber 2018. 

The data is downloaded as a zip folder with seperate CSV files for each month. As these are in individual folders the CSV need to be moved into the same folder. This was done manually in file explorer.

```{r data_load, echo = TRUE, eval = FALSE}
# Set directory to data folder 
setwd("~/R Programming/Coursera/Data Products/Shiny/LondonCrimeMap/data/MPS 2014-18")

# Create list of files to import
file_list <- list.files()

# Run function to import files
dataset <- do.call("rbind", lapply(file_list, read.csv, header = TRUE))

# Return directory to initial folder
setwd("~/R Programming/Coursera/Data Products/Shiny/LondonCrimeMap")

# Read in GIS data for mapping
boroughs<-readOGR(dsn="data/statistical-gis-boundaries-london/ESRI", layer="London_Borough_Excluding_MHW")

# Cut out unnecessary columns
boroughs@data<-boroughs@data[,c(1,2)]

# transform to WGS884 reference system 
boroughs<-spTransform(boroughs, CRS("+init=epsg:4326"))

# Find the edges of our map
bounds<-bbox(boroughs)

# Create a lits of boroughs for cleaning
borough_list <- boroughs@data[1]
```

## Preparing reporting datasets

The final dashboard will be comprised of three elements:

- Borough heatmap
- Interactive point map
- Crime trend graph

In order to facilitate this we will need to convert the raw data into two different dataframes, one with longitute and latitude per crime and one summarised by borough and crime type.

```{r summary_data}
 # Create subset for summary totals
crime_summary <- dataset[,c("Month", "LSOA.name", "Crime.type")] %>% 
  separate(LSOA.name, c("NAME", "Code"), sep = -5)

# Remove code variable
crime_summary$Code <- NULL

# Summarise data
crime_summary <- crime_summary %>%
  group_by(Month, NAME, Crime.type) %>%
  summarise(Total = n())

# Remove data from outside London
crime_summary <- crime_summary[crime_summary$NAME %in% borough_list$NAME,]

```


```{r location_data}
# Create subset with latitue and longitude data
crime_location <- dataset[,c("Month", "Longitude", "Latitude", "LSOA.name", "Crime.type", "Last.outcome.category")] %>% 
  separate(LSOA.name, c("NAME", "Code"), sep = -5)

# Remove code variable
crime_location$Code <- NULL

# Remove data from outside London
crime_location <- crime_location[crime_location$NAME %in% borough_list$NAME,]

```

